Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-11 PAPER.tex
made josh edits until 2.3

Because fringe-rate filtering is analogous to averaging in time, it comes at the cost of reducing the total number of independent 
samples in the data. To mimic this filter, we average every four time integrations of our toy model dataset together, yielding 
$25$ independent samples in time (Figure \ref{fig:toy_sigloss5}, left). We choose these numbers so that the total number of 
independent samples is similar to the number of frequency channels --- our matrices will still be full rank.
%\footnote{If instead we average every $20$ samples together, yielding only $5$ independent samples in time, we are left with a rank deficient covariance matrix. In this case... \cc{fill in}}

\cc{Note to Adrian - somewhere in the next paragraph would be good to ref Fig17}
The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure 
\ref{fig:toy_sigloss5} (right plot). The spectrum, in the case of fringe-rate filtering (dashed line), falls more steeply, especially for 
the last few eigenmodes. In many applications, this steep fall is not indicative of the true covariance structure of the data.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, 
yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\widehat{\textbf{C}}_{FG+EoR}$, 
in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green).}
	\label{fig:toy_sigloss5}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our `fringe-rate filtered' (time-averaged) toy model simulation --- foregrounds only (blue), 
EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each minimize signal 
loss, including modeling the covariance matrix of EoR (upper left), regularizing $\widehat{\textbf{C}}$ by adding an identity matrix to 
it (upper right), using only the first three eigenmodes of $\widehat{\textbf{C}}$ (lower left), and keeping only the diagonal elements of 
$\widehat{\textbf{C}}$ (lower right).}
	\label{fig:toy_sigloss8}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\widehat{\textbf{C}}$ (green) to that of four alternate 
weighting options, including modeling the covariance matrix of EoR (red), regularizing $\widehat{\textbf{C}}$ by adding an identity 
matrix to it (blue), using only the first three eigenmodes of $\widehat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
\widehat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
computed for the `fringe-rate filtered' (time-averaged) toy model case.}
	\label{fig:toy_sigloss14}
\end{figure}

@@

The last regularization scheme we are highlighting here is setting $\widehat{\textbf{C}} = \widehat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (keeping only the diagonal elements of $\widehat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we are only left with correlation information between the same 
two frequencies. Because we disregard information from all other frequency combinations in this case, we do a poor job 
suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.

Although the fourth method did not successfully recover EoR for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may be more effective than others based on the spectral nature 
of the components in a dataset. One may imagine a situation where a particular systematic is contained to an isolated 
frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
\widehat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

Going back to our toy model, we apply a sliding boxcar average to $10$ samples at a time, thus reducing the number of 
independent data samples to $N/10 = 100$. Bootstrapping this time-averaged noise, using the same method as described 
earlier (drawing $N$=$1000$ elements per bootstrap sample), under-estimates the error by a factor of $\sim3$ (black points in Figure \ref{fig:toy_error1}, at $N$=$1000$). This occurs 
because we are drawing more samples than independent ones available, and thus some samples are repeated multiple times 
in all bootstraps, leading to less variation between the bootstraps. In fact, the error derived from bootstrapping is a strong 
function of the number of elements that are drawn per bootstrap (Figure \ref{fig:toy_error1}, black points), and we can both 
under-estimate the error by drawing too many or over-estimate it by drawing too few. However, since we know that we have $100$ 
independent samples in this toy model, the error associated with drawing $N$=$100$ samples with replacement does match the theoretical prediction 
as expected.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation from bootstrapping as a function of the number of elements drawn per bootstrap when 
sampling with replacement. The star represents the standard deviation of $N_{boot}$ bootstraps, each created by drawing $1000$ 
elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be under-estimated if 
drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical 
prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

>>>>

Because fringe-rate filtering is analogous to averaging in time, it comes at the cost of reducing the total number of independent 
samples in the data. To mimic this filter, we average every four time integrations of our toy model dataset together, yielding 
$25$ independent samples in time (Figure \ref{fig:toy_sigloss5}). We choose these numbers so that the total number of 
independent samples is similar to the number of frequency channels --- our matrices will still be full rank.
%\footnote{If instead we average every $20$ samples together, yielding only $5$ independent samples in time, we are left with a rank deficient covariance matrix. In this case... \cc{fill in}}

\cc{Note to Adrian - somewhere in the next paragraph would be good to ref Fig17}
The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure 
\ref{fig:toy_sigloss15}. The spectrum, in the case of fringe-rate filtering (dashed line), falls more steeply, especially for 
the last few eigenmodes. In many applications, this steep fall is not indicative of the true covariance structure of the data.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss5.png}
	\caption{Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, 
yielding $25$ independent samples in time. Real parts are shown here.}
	\label{fig:toy_sigloss5}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.png}
	\caption{Eigenspectrum of $\widehat{\textbf{C}}_{FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green).}
	\label{fig:toy_sigloss15}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our `fringe-rate filtered' (time-averaged) toy model simulation --- foregrounds only (blue), 
EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each minimize signal 
loss, including modeling the covariance matrix of EoR (upper left), regularizing $\widehat{\textbf{C}}$ by adding an identity matrix to 
it (upper right), using only the first three eigenmodes of $\widehat{\textbf{C}}$ (lower left), and keeping only the diagonal elements of 
$\widehat{\textbf{C}}$ (lower right). The first case (upper left) is not feasible in practice since we do not know $\textbf{C}_{FG}$ and $\textbf{C}_{EoR}$ like we do in the toy model.}
	\label{fig:toy_sigloss8}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\widehat{\textbf{C}}$ (green) to that of four alternate 
weighting options, including modeling the covariance matrix of EoR (red), regularizing $\widehat{\textbf{C}}$ by adding an identity 
matrix to it (blue), using only the first three eigenmodes of $\widehat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
\widehat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
computed for the `fringe-rate filtered' (time-averaged) toy model case presented in Section \ref{sec:toymodel_frf}.}
	\label{fig:toy_sigloss14}
\end{figure}

@@

The last regularization scheme we are highlighting here is setting $\widehat{\textbf{C}} = \widehat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (keeping only the diagonal elements of $\widehat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does not down-weight the foregrounds. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we are only left with correlation information between the same 
two frequencies. Because we disregard information from all other frequency combinations in this case, we do a poor job 
suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss. Although this method did not successfully recover EoR for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may be more effective than others based on the spectral nature 
of the components in a dataset. 
%One may imagine a situation where a particular systematic is contained to an isolated 
%frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
%\widehat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

Going back to our toy model, we apply a sliding boxcar average to $10$ samples at a time, thus reducing the number of 
independent data samples to $N/10 = 100$. Bootstrapping this time-averaged noise, using the same method as described 
earlier (drawing $N$=$1000$ elements per bootstrap sample), under-estimates the error by a factor of $\sim3$ (black points in Figure \ref{fig:toy_error1}, at $N$=$1000$). This occurs 
because we are drawing more samples than independent ones available, and thus some samples are repeated multiple times 
in all bootstraps, leading to less variation between the bootstraps. In fact, the error derived from bootstrapping is a strong 
function of the number of elements that are drawn per bootstrap (Figure \ref{fig:toy_error1}, black points), and we can both 
under-estimate the error by drawing too many or over-estimate it by drawing too few. However, since we know that we have $100$ 
independent samples in this toy model, the error associated with drawing $N$=$100$ samples with replacement does match the theoretical prediction 
as expected (the black points cross the green line at $N$=$100$ in Figure \ref{fig:toy_error1}).

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation from bootstrapping as a function of the number of elements drawn per bootstrap when 
sampling with replacement. The star represents the standard deviation of $N_{boot}=500$ bootstraps, each created by drawing $1000$ 
elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be under-estimated if 
drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical 
prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}
